---
title: "Large Blank Correction for WS samples"
author: "Brett Longworth"
date: "4/21/2020"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---


```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

# Current LBC, why change?

* LBC for WS currently uses HY blanks
* WS blank is 4 times higher Fm than HY (USAMS)
  - HY approx 1 permil
  - WS approx 4 permil


```{r}
library(amstools)
library(tidyverse)
library(here)
library(knitr)
library(odbc)
library(blanks)

options(digits = 4)
theme_set(theme_light())

# Function to calculate SNICSer-style LBC error floors
Blankerr <- function(blankfm) {
  mfm <- mean(blankfm)
  sd <- sd(blankfm)
  ifelse(sd > mfm / 2, sd, mfm / 2)
}

data <- read_csv(here("data/wstd.csv"))

wstd <- data %>%
  filter(ws_method_num == 410305) %>% # only use WSL data
  mutate(Name = ordered(recode(as.character(rec_num), 
                                 "1081" = "C1", 
                                 "83028" = "C1", 
                                 "1082" = "C2", 
                                 "17185" = "TIRI-I",
                                 "159579" = "GW Blank")),
         fm_consensus = ifelse(rec_num == 159579, 0.00, fm_consensus)) %>%
  group_by(system, rec_num) %>%
  mutate(outlier = is.na(removeOutliers(norm_ratio))) 

# Calculate large blank correction by system
wstd <- wstd %>%
  group_by(system) %>%
      mutate(FmBlank = mean(norm_ratio[which(rec_num == 159579)]),
             FmBlankErr = Blankerr(norm_ratio[which(rec_num == 159579)]),
             FmLBC = doLBC(norm_ratio, FmBlank, 1.0398),
             FmLBCerr = doLBCerr(norm_ratio, FmBlank, 1.0398,
                                 pmax(int_err, ext_err), FmBlankErr))


# Calculate large blank correction by wheel
wstd <- wstd %>%
  group_by(wheel) %>%
      mutate(FmBlankWh = mean(norm_ratio[which(rec_num == 159579 & outlier == FALSE)]),
             FmBlankErrWh = Blankerr(norm_ratio[which(rec_num == 159579 & outlier == FALSE)]), 
             NblankWh = length(norm_ratio[which(rec_num == 159579 & outlier == FALSE)]),
             FmLBCWh = ifelse(NblankWh > 1, doLBC(norm_ratio, FmBlank, 1.0398), NA), # only calc for wheels with 2 blanks
             FmLBCerrWh = ifelse(NblankWh > 1, doLBCerr(norm_ratio, FmBlank, 1.0398,
                                 pmax(int_err, ext_err), FmBlankErr), NA))
```

---
# Definition of the large blank Fm

* Fm of Ft. Caswell groundwater and dissolved C-1 are comparable
* Can be used interchangably as LBC blank
* Several C-1 outliers removed

```{r}
blanks <- wstd %>% 
	filter(rec_num %in% c(83028, 159579)) 

blanks %>%
  group_by(Name, system) %>%
  summarize(Fm = mean(norm_ratio),
            "Fm SD" = sd(norm_ratio),
            "Fm (no outliers)" = mean(norm_ratio[which(outlier == FALSE & norm_ratio < 0.02)]),
            "Fm SD (no outliers)" = sd(norm_ratio[which(outlier == FALSE & norm_ratio < 0.02)]),
            N = n()) %>%
  arrange(Name) %>%
  kable('html')
```

---

# Distribution of WS blanks

```{r}
blanks %>% 
	filter(rec_num %in% c(83028, 159579)) %>%
  ggplot(aes(norm_ratio, fill = system)) +
    geom_histogram(alpha = .8) +
    scale_color_viridis_d() +
    facet_grid(rows = vars(Name), scales = "fixed") +
    labs(title = "Normalized Fm of WS blanks",
         x = "Normalized Fm (no LBC)")
```

---

# Without outliers


```{r}
wstd %>% 
	filter(rec_num %in% c(83028, 159579),
	       outlier == FALSE,
	       norm_ratio < 0.02) %>%
  ggplot(aes(norm_ratio, fill = system)) +
    geom_histogram(alpha = .8) +
    scale_color_viridis_d() +
    facet_grid(rows = vars(Name), scales = "fixed") +
    labs(title = "Normalized Fm of WS blanks, outliers removed",
         x = "Normalized Fm (no LBC)")

```

---

# Results of the correction for dissolved secondaries


```{r}
wstd%>% 
      filter(!is.na(FmLBC),
             !is.na(fm_consensus)) %>%
  group_by(Name, rec_num, system) %>%
        summarize("No Blank" = mean(norm_ratio), 
                  "HY LBC" = mean(fm_corr),
                  "HY LBC err" = mean(sig_fm_corr),
                  "WS LBC" = mean(FmLBC),
                  "WS LBC err" = mean(FmLBCerr),
                  fm_consensus = mean(fm_consensus),
                  N = n()) %>%
        arrange(rec_num) %>%
  kable('html')
```

---
# Correction results

```{r}
wstd %>%
  filter(!is.na(fm_consensus),
         outlier == FALSE,
         !(rec_num == 83028 & norm_ratio > 0.02)) %>%
  rename("No LBC" = norm_ratio, "HY LBC" = fm_corr, "WS LBC" = FmLBC) %>%
  pivot_longer(c("No LBC", "HY LBC", "WS LBC"), names_to = "Method", values_to = "Fm") %>%
  ggplot(aes(fct_reorder(Method, Fm, .desc = TRUE), Fm)) +
  geom_hline(aes(yintercept = fm_consensus) ) +
  geom_boxplot() +
  geom_jitter() +
  facet_grid(rows = vars(Name), cols = vars(system), scales = "free") +
  labs(title = "Distribution of HY and WS corrected Fm for secondary standards and blanks",
       x = "Blank cor method")
  

```

---
# Modeling the correction for varying sample Fm

* Modelled effect of LBC on ratio for HY blank and WS blank
* Difference is ~3 permil for dead samples, ~0.5 permil for nearly-modern


---
# Modeling the correction 

```{r}
# curve function
fun <- function(Fm, FmB) {
  bcfm <- doLBC(Fm, FmB, 1.0398)
  Fm - bcfm
}

# model plot
base <- ggplot(data.frame(Fm = c(0, 1.5)), aes(Fm))
base + 
  geom_hline(yintercept = 0) +
  stat_function(fun = fun, args = list(FmB = 0.004), aes(color = "WS LBC"), size = 1) +
  stat_function(fun = fun, args = list(FmB = 0.001), aes(color = "HY LBC"), size = 1) +
  scale_color_manual("LBC curve", values = c("orange", "green")) +
  labs(title = "Effect of LBC on Fm for HY vs WS blank",
       color = "Blank value",
       y = "Fm difference (meas - LBC)")
  
```


---

# Model with secondary and blank data

```{r}
wstd %>%
  filter(!is.na(fm_consensus),
         outlier == FALSE,
         system == "USAMS",
         !(rec_num == 83028 & norm_ratio > 0.02)) %>%
  mutate(Fm = norm_ratio,
         Fm_diff = norm_ratio - fm_consensus) %>%
  ggplot(aes(Fm, Fm_diff, shape = Name)) +
  geom_hline(yintercept = 0) +
  stat_function(fun = fun, args = list(FmB = 0.004), aes(color = "WS LBC"), size = 1) +
  stat_function(fun = fun, args = list(FmB = 0.001), aes(color = "HY LBC"), size = 1) +
  geom_point() +
  xlim(0,1.5) +
  scale_color_manual("LBC curve", values = c("orange", "green")) +
  labs(title = "Data difference from consensus with HY vs WS blank curves",
       color = "Blank value",
       y = "Fm difference (meas - LBC)")
  
  
```

---
# Mass dependence

* C-1 data may show mass-dependence
* not enough data for other sample types
* dillution experiment with groundwater blank and modern seawater standard may provide adequate data for curves


---
# Mass dependence

```{r}
wstd %>%
  filter(!is.na(fm_consensus),
         outlier == FALSE,
         rec_num == 83028,
         norm_ratio < 0.02) %>%
  ggplot(aes(total_umols_co2, norm_ratio)) + 
  geom_point()
```

# Next steps

* Using the water blanks to perform a large blank correction has a significant effect on reported ratios that differ from the ratio of the standard. 
* This correction would improve the accuracy of water sample measurements, especially at low Fm. 
* Developing a constant contamination/mass balance model
  - Samples of differing mass would get an appropriate correction
  - Large mass samples would get a correction that accounts for the fact that the normalizing standards do not contain the process blank for the water lines.

