---
title: "DIC Blanks"
author: "Brett Longworth"
date: "3/12/2020"
output:
  html_document: default
  word_document: default
  pdf_document: default
---

What is the effect of using WS process blanks for correcting water secondaries? 

1. Extract DIC data by wheel.
2. Find WS blanks by wheel and average
3. Apply LBC using WS blanks
4. Compare with DIC corrected with HY blanks

TODO: 
Separate REDICS data 
Need process or reactor number

look at seawater standards
Get data for ARM, etc...

remove outliers
Use 1.5x interquartile

estimate effect of the correction on GO SHIP and USGS data.
Maybe just plot average correction amount vs Fm

Apply correction to REDICS using REDICS blanks and check seawater standard agreement


Fort Caswell groundwater blanks have roughly the same value as C1 WS samples, and there are many more of them, so I used them for blank correction.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

## WS standards overview

A summary of the samples used here. Data for some sample types have a lot of spread. I haven't run the analysis after removing outlers yet.

```{r}
library(amstools)
library(tidyverse)
library(here)
library(knitr)
library(odbc)
library(blanks)

options(digits = 5)
theme_set(theme_light())

# Function to calculate SNICSer-style LBC error floors
Blankerr <- function(blankfm) {
  mfm <- mean(blankfm)
  sd <- sd(blankfm)
  ifelse(sd > mfm / 2, sd, mfm / 2)
}

data <- read_csv(here("data/wstd.csv"))

wstd <- data %>%
  filter(ws_method_num == 410305) %>% # only use WSL data
  mutate(Name = ordered(recode(as.character(rec_num), 
                                 "1081" = "C1", 
                                 "83028" = "C1", 
                                 "1082" = "C2", 
                                 "17185" = "TIRI-I",
                                 "159579" = "GW Blank")),
         fm_consensus = ifelse(rec_num == 159579, 0.00, fm_consensus)) 
```

Data summary

```{r}
wstd %>% 
        group_by(Name, ws_method_num, system) %>%
        summarize(Fm = mean(fm_corr),
                  "Fm SD" = sd(fm_corr),
                  N = n()) %>%
        arrange(Name) %>%
  kable

ggplot(wstd, aes(norm_ratio, fill = system)) +
  geom_histogram(alpha = .8) +
  scale_color_viridis_d() +
  facet_grid(cols = vars(Name), scales = "free")
```

## Correction using mean system water blank

Because not all wheels with DIC secondaries had two WS blanks, we can use a system average blank value to apply the large blank correction by system. The blank corrected errors will be a little low, because I haven't applied SNICSer's "error floor" method yet.

```{r}
wstdsys <- wstd %>%
  group_by(system) %>%
      mutate(FmBlank = mean(norm_ratio[which(rec_num == 159579)]),
             FmBlankErr = Blankerr(norm_ratio[which(rec_num == 159579)]), #use SNICSer error calc here
             FmLBC = doLBC(norm_ratio, FmBlank, 1.0398),
             FmLBCerr = doLBCerr(norm_ratio, FmBlank, 1.0398,
                                 pmax(int_err, ext_err), FmBlankErr))


wstdsys %>% 
        group_by(Name, rec_num, system) %>%
        summarize("No Blank" = mean(norm_ratio), 
                  "HY LBC" = mean(fm_corr),
                  "HY LBC err" = mean(sig_fm_corr),
                  "WS LBC" = mean(FmLBC),
                  "WS LBC err" = mean(FmLBCerr),
                  fm_consensus = mean(fm_consensus),
                  N = n()) %>%
        arrange(rec_num) %>%
  kable()

```

```{r}
wstdsys %>%
  filter(!is.na(FmLBC)) %>%
  pivot_longer(c(fm_corr, FmLBC), names_to = "Method", values_to = "Fm") %>%
  ggplot(aes(Method, Fm)) +
  geom_hline(aes(yintercept = fm_consensus) ) +
  #geom_hline(yintercept = 0.2524) +
  #geom_hline(yintercept = 0.4114) +
  geom_boxplot() +
  facet_grid(rows = vars(Name), cols = vars(system), scales = "free")
```

The same with outliers removed

```{r}
wstdsys <- wstd %>%
  group_by(Name, system) %>%
  filter(!is.na(amstools::removeOutliers(norm_ratio))) %>%
  ungroup() %>%
  group_by(system) %>%
      mutate(FmBlank = mean(norm_ratio[which(rec_num == 159579)]),
             FmBlankErr = Blankerr(norm_ratio[which(rec_num == 159579)]), 
             FmLBC = doLBC(norm_ratio, FmBlank, 1.0398),
             FmLBCerr = doLBCerr(norm_ratio, FmBlank, 1.0398,
                                 pmax(int_err, ext_err), FmBlankErr))


wstdsys %>% 
        group_by(Name, rec_num, system) %>%
        summarize("No Blank" = mean(norm_ratio), 
                  "HY LBC" = mean(fm_corr),
                  "HY LBC err" = mean(sig_fm_corr),
                  "WS LBC" = mean(FmLBC),
                  "WS LBC err" = mean(FmLBCerr),
                  fm_consensus = mean(fm_consensus),
                  N = n()) %>%
        arrange(rec_num) %>%
  kable()

```

```{r}
wstdsys %>%
  filter(!is.na(FmLBC)) %>%
  pivot_longer(c(fm_corr, FmLBC), names_to = "Method", values_to = "Fm") %>%
  ggplot(aes(Method, Fm)) +
  geom_hline(aes(yintercept = fm_consensus) ) +
  #geom_hline(yintercept = 0.2524) +
  #geom_hline(yintercept = 0.4114) +
  geom_boxplot() +
  facet_grid(rows = vars(Name), cols = vars(system), scales = "free")
```

## Correction using per-wheel water blank

For wheels with two or more water blanks, we can calculate and apply a large blank correction based on the per-wheel blank.

```{r}
wstdwheel <- wstd %>%
  group_by(wheel) %>%
      mutate(FmBlank = mean(norm_ratio[which(rec_num == 159579)]),
             FmBlankErr = Blankerr(norm_ratio[which(rec_num == 159579)]), 
             Nblank = length(norm_ratio[which(rec_num == 159579)]),
             FmLBC = ifelse(Nblank > 1, doLBC(norm_ratio, FmBlank, 1.0398), NA),
             FmLBCerr = ifelse(Nblank > 1, doLBCerr(norm_ratio, FmBlank, 1.0398,
                                 pmax(int_err, ext_err), FmBlankErr), NA))

wstdwheel %>% 
      filter(!is.na(FmLBC)) %>%
  group_by(Name, rec_num, system) %>%
        summarize("No Blank" = mean(norm_ratio), 
                  "HY LBC" = mean(fm_corr),
                  "HY LBC err" = mean(sig_fm_corr),
                  "WS LBC" = mean(FmLBC),
                  "WS LBC err" = mean(FmLBCerr),
                  fm_consensus = mean(fm_consensus),
                  N = n()) %>%
        arrange(rec_num) %>%
  kable()

```

```{r}
wstdwheel %>%
  filter(!is.na(FmLBC)) %>%
  pivot_longer(c(fm_corr, FmLBC), names_to = "Method", values_to = "Fm") %>%
  ggplot(aes(Method, Fm)) +
  geom_hline(aes(yintercept = fm_consensus) ) +
  #geom_hline(yintercept = 0.2524) +
  #geom_hline(yintercept = 0.4114) +
  geom_boxplot() +
  facet_grid(rows = vars(Name), cols = vars(system), scales = "free")
```


The same without outliers

```{r}
wstdwheel <- wstd %>%
  group_by(Name, system) %>%
  filter(!is.na(amstools::removeOutliers(norm_ratio))) %>%
  ungroup() %>%
  group_by(wheel) %>%
      mutate(FmBlank = mean(norm_ratio[which(rec_num == 159579)]),
             FmBlankErr = Blankerr(norm_ratio[which(rec_num == 159579)]), 
             Nblank = length(norm_ratio[which(rec_num == 159579)]),
             FmLBC = ifelse(Nblank > 1, doLBC(norm_ratio, FmBlank, 1.0398), NA),
             FmLBCerr = ifelse(Nblank > 1, doLBCerr(norm_ratio, FmBlank, 1.0398,
                                 pmax(int_err, ext_err), FmBlankErr), NA))

wstdwheel %>% 
      filter(!is.na(FmLBC)) %>%
  group_by(Name, rec_num, system) %>%
        summarize("No Blank" = mean(norm_ratio), 
                  "HY LBC" = mean(fm_corr),
                  "HY LBC err" = mean(sig_fm_corr),
                  "WS LBC" = mean(FmLBC),
                  "WS LBC err" = mean(FmLBCerr),
                  fm_consensus = mean(fm_consensus),
                  N = n()) %>%
        arrange(rec_num) %>%
  kable()

```

```{r}
wstdwheel %>%
  filter(!is.na(FmLBC)) %>%
  pivot_longer(c(fm_corr, FmLBC), names_to = "Method", values_to = "Fm") %>%
  ggplot(aes(Method, Fm)) +
  geom_hline(aes(yintercept = fm_consensus) ) +
  #geom_hline(yintercept = 0.2524) +
  #geom_hline(yintercept = 0.4114) +
  geom_boxplot() +
  facet_grid(rows = vars(Name), cols = vars(system), scales = "free")
```
