---
title: "DIC Blanks"
author: "Brett Longworth"
date: "3/12/2020"
output:
  html_document: default
  word_document: default
  pdf_document: default
---

TODO: 

look at seawater standards
Get data for ARM, etc...

remove outliers
Use 3x interquartile

estimate effect of the correction on GO SHIP and USGS data.
Maybe just plot average correction amount vs Fm

Apply correction to REDICS using REDICS blanks and check seawater standard agreement
Currently, the large blank correction applied to the WS sample type is based on
the HY blanks on a given wheel. Data from dissolved C-1 blank material and
radiocarbon-dead water from the Ft. Caswell groundwater site show that large,
dead samples return values around Fm 0.003, indicating that the process has a
significant non-dead blank component. In this document, I define this blank and
show the effect of using it to correct dissolved secondary standards and model
the effect on large samples of varying Fm. 

A constant contamination or mass balance correction would be required to fully
correct for process blank contribution for variable sample sizes. We do not
have the neccessary data to determine the mass and fraction modern of the blank
at this point. 

```{r}
library(amstools)
library(tidyverse)
library(here)
library(knitr)
library(odbc)
library(blanks)

options(digits = 5)
theme_set(theme_light())

# Function to calculate SNICSer-style LBC error floors
Blankerr <- function(blankfm) {
  mfm <- mean(blankfm)
  sd <- sd(blankfm)
  ifelse(sd > mfm / 2, sd, mfm / 2)
}

data <- read_csv(here("data/wstd.csv"))

wstd <- data %>%
  filter(ws_method_num == 410305) %>% # only use WSL data
  mutate(Name = ordered(recode(as.character(rec_num), 
                                 "1081" = "C1", 
                                 "83028" = "C1", 
                                 "1082" = "C2", 
                                 "17185" = "TIRI-I",
                                 "159579" = "GW Blank")),
         fm_consensus = ifelse(rec_num == 159579, 0.00, fm_consensus)) %>%
  group_by(system, rec_num) %>%
  mutate(outlier = is.na(removeOutliers(norm_ratio))) 

# Calculate large blank correction by system
wstd <- wstd %>%
  group_by(system) %>%
      mutate(FmBlank = mean(norm_ratio[which(rec_num == 159579)]),
             FmBlankErr = Blankerr(norm_ratio[which(rec_num == 159579)]),
             FmLBC = doLBC(norm_ratio, FmBlank, 1.0398),
             FmLBCerr = doLBCerr(norm_ratio, FmBlank, 1.0398,
                                 pmax(int_err, ext_err), FmBlankErr))


# Calculate large blank correction by wheel
wstd <- wstd %>%
  group_by(wheel) %>%
      mutate(FmBlankWh = mean(norm_ratio[which(rec_num == 159579)]),
             FmBlankErrWh = Blankerr(norm_ratio[which(rec_num == 159579)]), 
             NblankWh = length(norm_ratio[which(rec_num == 159579)]),
             FmLBCWh = ifelse(Nblank > 1, doLBC(norm_ratio, FmBlank, 1.0398), NA), # only calc for wheels with 2 blanks
             FmLBCerrWh = ifelse(Nblank > 1, doLBCerr(norm_ratio, FmBlank, 1.0398,
                                 pmax(int_err, ext_err), FmBlankErr), NA))
```


# Definition of the large blank Fm

The table below shows data for C-1 and Ft. Caswell groundwater. The data show that the Fm of these sample types is comparable, meaning that both can be used as a representative process blank for WS. The blank for REDICS and the WSL are slightly different, therefore we should consider applying a different large blank for these processes.

```{r}
wstd %>% 
	filter(rec_num %in% c(1081, 159579) %>%
        group_by(Name, ws_method_num, system) %>%
        summarize(Fm = mean(fm_corr),
                  "Fm SD" = sd(fm_corr),
                  N = n()) %>%
        arrange(Name) %>%
  kable

ggplot(wstd, aes(norm_ratio, fill = system)) +
  geom_histogram(alpha = .8) +
  scale_color_viridis_d() +
  facet_grid(cols = vars(Name), scales = "free")
```


# Results of the correction for dissolved secondaries

We can now use the Fm value of the water blanks to apply a large blank correction to water samples. This is applied using the LBC and LBC error formulas from SNICSer. I've looked at applying these using a per-wheel blank Fm average and a per-system average. Because each wheel with water samples didn't include two water blanks, data from the per-wheel method is sparse.


```{r}
wstdwheel %>% 
      filter(!is.na(FmLBC)) %>%
  group_by(Name, rec_num, system) %>%
        summarize("No Blank" = mean(norm_ratio), 
                  "HY LBC" = mean(fm_corr),
                  "HY LBC err" = mean(sig_fm_corr),
                  "WS LBC" = mean(FmLBC),
                  "WS LBC err" = mean(FmLBCerr),
                  fm_consensus = mean(fm_consensus),
                  N = n()) %>%
        arrange(rec_num) %>%
  kable()
```


```{r}
wstdwheel %>%
  filter(!is.na(FmLBC)) %>%
  pivot_longer(c(fm_corr, FmLBC), names_to = "Method", values_to = "Fm") %>%
  ggplot(aes(Method, Fm)) +
  geom_hline(aes(yintercept = fm_consensus) ) +
  #geom_hline(yintercept = 0.2524) +
  #geom_hline(yintercept = 0.4114) +
  geom_boxplot() +
  facet_grid(rows = vars(Name), cols = vars(system), scales = "free")
```


# Modeling the correction for varying sample Fm

Given the large blank correction formula and knowing the fraction modern of the water blank samples, we can model the effect of the correction of samples by their Fm. 

TODO: model blank. Fm change vs Fm. Should be ~3permil @ Fm 0, 0 correction at Fm 1.0398


# Next steps

It is clear that using the water blanks to perform a large blank correction has a significant effect on reported ratios that differ from the ratio of the standard. This correction alone would improve the accuracy of water sample measurements, especially at low Fm. 

Developing a constant contamination/mass balance model for water samples would be useful for two reasons. First, samples of differing mass would get an appropriate correction. Second, large mass samples would get a correction that accounts for the fact that the normalizing standards do not contain the process blank for the water lines.

Samples run on REDICS show that this system has a smaller blank than the WSL. The large and mass balance blank of both lines should be compared to determine whether separate blanks should be used for each line.









