---
title: "DIC Blanks"
author: "Brett Longworth"
date: "3/12/2020"
output:
  html_document: default
  word_document: default
  pdf_document: default
---

Currently, the large blank correction applied to the WS sample type is based on the HY blanks on a given wheel. Data from dissolved C-1 blank material and radiocarbon-dead water from the Ft. Caswell groundwater site show that large, dead samples return values around Fm 0.003, indicating that the process has a significant non-dead blank component. In this document, I define this blank and show the effect of using it to correct dissolved secondary standards and model the effect on large samples of varying Fm. 

A constant contamination or mass balance correction would be required to fully correct for process blank contribution for variable sample sizes. We do not have the neccessary data to determine the mass and fraction modern of the blank at this point. 

# Definition of the large blank Fm

# Results of the correction for dissolved secondaries

# Modeling the correction for varying sample Fm

# Next steps

What is the effect of using WS process blanks for correcting water secondaries? 

1. Extract DIC data by wheel.
2. Find WS blanks by wheel and average
3. Apply LBC using WS blanks
4. Compare with DIC corrected with HY blanks

TODO: 
Separate REDICS data 
Need process or reactor number

look at seawater standards
Get data for ARM, etc...

remove outliers
Use 3x interquartile

estimate effect of the correction on GO SHIP and USGS data.
Maybe just plot average correction amount vs Fm

Apply correction to REDICS using REDICS blanks and check seawater standard agreement


Fort Caswell groundwater blanks have roughly the same value as C1 WS samples, and there are many more of them, so I used them for blank correction.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

## WS standards overview

A summary of the samples used here. Data for some sample types have a lot of spread. I haven't run the analysis after removing outlers yet.

```{r}
library(amstools)
library(tidyverse)
library(here)
library(knitr)
library(odbc)
library(blanks)

options(digits = 5)
theme_set(theme_light())

# Function to calculate SNICSer-style LBC error floors
Blankerr <- function(blankfm) {
  mfm <- mean(blankfm)
  sd <- sd(blankfm)
  ifelse(sd > mfm / 2, sd, mfm / 2)
}

data <- read_csv(here("data/wstd.csv"))

wstd <- data %>%
  filter(ws_method_num == 410305) %>% # only use WSL data
  mutate(Name = ordered(recode(as.character(rec_num), 
                                 "1081" = "C1", 
                                 "83028" = "C1", 
                                 "1082" = "C2", 
                                 "17185" = "TIRI-I",
                                 "159579" = "GW Blank")),
         fm_consensus = ifelse(rec_num == 159579, 0.00, fm_consensus)) %>%
  group_by(system, rec_num) %>%
  mutate(outlier = is.na(removeOutliers(norm_ratio))) 
```

Data summary

```{r}
wstd %>% 
        group_by(Name, ws_method_num, system) %>%
        summarize(Fm = mean(fm_corr),
                  "Fm SD" = sd(fm_corr),
                  N = n()) %>%
        arrange(Name) %>%
  kable

ggplot(wstd, aes(norm_ratio, fill = system)) +
  geom_histogram(alpha = .8) +
  scale_color_viridis_d() +
  facet_grid(cols = vars(Name), scales = "free")
```

## Correction using mean system water blank

Because not all wheels with DIC secondaries had two WS blanks, we can use a system average blank value to apply the large blank correction by system. The blank corrected errors will be a little low, because I haven't applied SNICSer's "error floor" method yet.

```{r}
wstdsys <- wstd %>%
  group_by(system) %>%
      mutate(FmBlank = mean(norm_ratio[which(rec_num == 159579)]),
             FmBlankErr = Blankerr(norm_ratio[which(rec_num == 159579)]),
             FmLBC = doLBC(norm_ratio, FmBlank, 1.0398),
             FmLBCerr = doLBCerr(norm_ratio, FmBlank, 1.0398,
                                 pmax(int_err, ext_err), FmBlankErr))


wstdsys %>% 
        group_by(Name, rec_num, system) %>%
        summarize("No Blank" = mean(norm_ratio), 
                  "HY LBC" = mean(fm_corr),
                  "HY LBC err" = mean(sig_fm_corr),
                  "WS LBC" = mean(FmLBC),
                  "WS LBC err" = mean(FmLBCerr),
                  fm_consensus = mean(fm_consensus),
                  N = n()) %>%
        arrange(rec_num) %>%
  kable()

```

```{r}
wstdsys %>%
  filter(!is.na(FmLBC)) %>%
  pivot_longer(c(fm_corr, FmLBC), names_to = "Method", values_to = "Fm") %>%
  ggplot(aes(Method, Fm, color = outlier)) +
  geom_hline(aes(yintercept = fm_consensus) ) +
  #geom_hline(yintercept = 0.2524) +
  #geom_hline(yintercept = 0.4114) +
  geom_boxplot() +
  geom_jitter() +
  facet_grid(rows = vars(Name), cols = vars(system), scales = "free")
```

The same with outliers removed

```{r}
wstdsys <- wstd %>%
  group_by(Name, system) %>%
  filter(!is.na(amstools::removeOutliers(norm_ratio))) %>%
  ungroup() %>%
  group_by(system) %>%
      mutate(FmBlank = mean(norm_ratio[which(rec_num == 159579)]),
             FmBlankErr = Blankerr(norm_ratio[which(rec_num == 159579)]), 
             FmLBC = doLBC(norm_ratio, FmBlank, 1.0398),
             FmLBCerr = doLBCerr(norm_ratio, FmBlank, 1.0398,
                                 pmax(int_err, ext_err), FmBlankErr))


wstdsys %>% 
        group_by(Name, rec_num, system) %>%
        summarize("No Blank" = mean(norm_ratio), 
                  "HY LBC" = mean(fm_corr),
                  "HY LBC err" = mean(sig_fm_corr),
                  "WS LBC" = mean(FmLBC),
                  "WS LBC err" = mean(FmLBCerr),
                  fm_consensus = mean(fm_consensus),
                  N = n()) %>%
        arrange(rec_num) %>%
  kable()

```

```{r}
wstdsys %>%
  filter(!is.na(FmLBC)) %>%
  pivot_longer(c(fm_corr, FmLBC), names_to = "Method", values_to = "Fm") %>%
  ggplot(aes(Method, Fm)) +
  geom_hline(aes(yintercept = fm_consensus) ) +
  #geom_hline(yintercept = 0.2524) +
  #geom_hline(yintercept = 0.4114) +
  geom_boxplot() +
  facet_grid(rows = vars(Name), cols = vars(system), scales = "free")
```

## Correction using per-wheel water blank

For wheels with two or more water blanks, we can calculate and apply a large blank correction based on the per-wheel blank.

```{r}
wstdwheel <- wstd %>%
  group_by(wheel) %>%
      mutate(FmBlank = mean(norm_ratio[which(rec_num == 159579)]),
             FmBlankErr = Blankerr(norm_ratio[which(rec_num == 159579)]), 
             Nblank = length(norm_ratio[which(rec_num == 159579)]),
             FmLBC = ifelse(Nblank > 1, doLBC(norm_ratio, FmBlank, 1.0398), NA),
             FmLBCerr = ifelse(Nblank > 1, doLBCerr(norm_ratio, FmBlank, 1.0398,
                                 pmax(int_err, ext_err), FmBlankErr), NA))

wstdwheel %>% 
      filter(!is.na(FmLBC)) %>%
  group_by(Name, rec_num, system) %>%
        summarize("No Blank" = mean(norm_ratio), 
                  "HY LBC" = mean(fm_corr),
                  "HY LBC err" = mean(sig_fm_corr),
                  "WS LBC" = mean(FmLBC),
                  "WS LBC err" = mean(FmLBCerr),
                  fm_consensus = mean(fm_consensus),
                  N = n()) %>%
        arrange(rec_num) %>%
  kable()

```

```{r}
wstdwheel %>%
  filter(!is.na(FmLBC)) %>%
  pivot_longer(c(fm_corr, FmLBC), names_to = "Method", values_to = "Fm") %>%
  ggplot(aes(Method, Fm)) +
  geom_hline(aes(yintercept = fm_consensus) ) +
  #geom_hline(yintercept = 0.2524) +
  #geom_hline(yintercept = 0.4114) +
  geom_boxplot() +
  facet_grid(rows = vars(Name), cols = vars(system), scales = "free")
```


The same without outliers

```{r}
wstdwheel <- wstd %>%
  group_by(Name, system) %>%
  filter(!is.na(amstools::removeOutliers(norm_ratio))) %>%
  ungroup() %>%
  group_by(wheel) %>%
      mutate(FmBlank = mean(norm_ratio[which(rec_num == 159579)]),
             FmBlankErr = Blankerr(norm_ratio[which(rec_num == 159579)]), 
             Nblank = length(norm_ratio[which(rec_num == 159579)]),
             FmLBC = ifelse(Nblank > 1, doLBC(norm_ratio, FmBlank, 1.0398), NA),
             FmLBCerr = ifelse(Nblank > 1, doLBCerr(norm_ratio, FmBlank, 1.0398,
                                 pmax(int_err, ext_err), FmBlankErr), NA))

wstdwheel %>% 
      filter(!is.na(FmLBC)) %>%
  group_by(Name, rec_num, system) %>%
        summarize("No Blank" = mean(norm_ratio), 
                  "HY LBC" = mean(fm_corr),
                  "HY LBC err" = mean(sig_fm_corr),
                  "WS LBC" = mean(FmLBC),
                  "WS LBC err" = mean(FmLBCerr),
                  fm_consensus = mean(fm_consensus),
                  N = n()) %>%
        arrange(rec_num) %>%
  kable()

```

```{r}
wstdwheel %>%
  filter(!is.na(FmLBC)) %>%
  pivot_longer(c(fm_corr, FmLBC), names_to = "Method", values_to = "Fm") %>%
  ggplot(aes(Method, Fm)) +
  geom_hline(aes(yintercept = fm_consensus) ) +
  #geom_hline(yintercept = 0.2524) +
  #geom_hline(yintercept = 0.4114) +
  geom_boxplot() +
  facet_grid(rows = vars(Name), cols = vars(system), scales = "free")
```
