---
title: "DIC Blanks"
author: "Brett Longworth"
date: "Last compiled on `r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document: default
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

Currently, the large blank correction applied to the WS sample type is based on
the HY blanks on a given wheel. Data from dissolved C-1 blank material and
radiocarbon-dead water from the Ft. Caswell groundwater site show that large,
dead samples return values around Fm 0.004, indicating that the process has a
significant non-dead blank component. In this document, I define this blank and
show the effect of using it to correct dissolved secondary standards and model
the effect on large samples of varying Fm. 

This analysis considers only WSL data. REDICS blank data differ, and we should consider whether a separate LBC for REDICS is necessary.

```{r}
library(amstools)
library(tidyverse)
library(here)
library(knitr)
library(odbc)
library(tidymodels)
#library(blanks)

options(digits = 4)
theme_set(theme_light())

# Function to calculate SNICSer-style LBC error floors
Blankerr <- function(blankfm) {
  mfm <- mean(blankfm)
  sd <- sd(blankfm)
  ifelse(sd > mfm / 2, sd, mfm / 2)
}

data <- read_csv(here("data/wstd.csv"))

wstd <- data %>%
  filter(ws_method_num == 410305, # only use WSL data
         total_umols_co2 * 12.015 > 900) %>% # only large samples 
  mutate(Name = ordered(recode(as.character(rec_num), 
                                 "1081" = "C1", 
                                 "83028" = "C1", 
                                 "1082" = "C2", 
                                 "2138" = "TIRI-F",
                                 "17185" = "TIRI-I",
                                 "38809" = "NOSAMS2",
                                 "159579" = "GW Blank")),
         fm_consensus = ifelse(rec_num == 159579, 0.00, fm_consensus)) %>%
  group_by(system, rec_num) %>%
  mutate(outlier = is.na(removeOutliers(norm_ratio))) %>% 
  ungroup()

# Calculate large blank correction by system
wstd <- wstd %>%
  group_by(system) %>%
      mutate(FmBlank = mean(norm_ratio[which(rec_num == 159579 & 
                                             outlier == FALSE & norm_ratio < 0.015)]),
             FmBlankErr = Blankerr(norm_ratio[which(rec_num == 159579 &
                                                    outlier == FALSE & norm_ratio < 0.015)]),
             FmLBC = doLBC(norm_ratio, FmBlank, 1.0398),
             FmLBCerr = doLBCerr(norm_ratio, FmBlank, 1.0398,
                                 pmax(int_err, ext_err), FmBlankErr)) %>% 
  ungroup()


# Calculate large blank correction by wheel
wstd <- wstd %>%
  group_by(wheel) %>%
      mutate(FmBlankWh = mean(norm_ratio[which(rec_num == 159579 & outlier == FALSE)]),
             FmBlankErrWh = Blankerr(norm_ratio[which(rec_num == 159579 & outlier == FALSE)]), 
             NblankWh = length(norm_ratio[which(rec_num == 159579 & outlier == FALSE)]),
             FmLBCWh = ifelse(NblankWh > 1, doLBC(norm_ratio, FmBlank, 1.0398), NA), # only calc for wheels with 2 blanks
             FmLBCerrWh = ifelse(NblankWh > 1, doLBCerr(norm_ratio, FmBlank, 1.0398,
                                 pmax(int_err, ext_err), FmBlankErr), NA)) %>% 
  ungroup()
```


# Definition of the large blank Fm

The table and plots below show data for C-1 and Ft. Caswell groundwater. The data show that the Fm of these sample types is comparable, meaning that both can be used as a representative process blank for WS. There have been intermittent problems with the dissolution of C-1 which caused high values that are removed in the "no outliers" columns and in the second plot below. Points > 1.5% modern or more than 1.5 * IQR above the 3rd quartile are considered outliers. These outlier values are not used for LBC.

For comparison the mean normalized ratio of the TIRI-F HY blank for the last 6 months was 0.0038 for CFAMS.

```{r}
blanks <- wstd %>% 
	filter(rec_num %in% c(1081, 2138, 83028, 159579)) 

blanks %>%
  group_by(Name, system) %>%
  summarize(Fm = mean(norm_ratio),
            "Fm SD" = sd(norm_ratio),
            "Fm (no outliers)" = mean(norm_ratio[which(outlier == FALSE & norm_ratio < 0.015)]),
            "Fm SD (no outliers)" = sd(norm_ratio[which(outlier == FALSE & norm_ratio < 0.015)]),
            N = n()) %>%
  arrange(Name) %>%
  kable

blanks %>% 
	filter(rec_num %in% c(1081, 2138, 83028, 159579)) %>%
  ggplot(aes(norm_ratio, fill = system)) +
    geom_histogram(alpha = .5, position = "identity") +
    scale_color_viridis_d() +
    facet_grid(rows = vars(Name), scales = "fixed") +
    labs(title = "Normalized Fm of WS blanks",
         x = "Normalized Fm (no LBC)")
```

After removing outliers.

```{r}
wstd %>% 
	filter(rec_num %in% c(1081, 2138, 83028, 159579),
	       outlier == FALSE,
	       norm_ratio < 0.015) %>%
  ggplot(aes(norm_ratio, fill = system)) +
    geom_histogram(alpha = .5, position = "identity") +
    scale_color_viridis_d() +
    facet_grid(rows = vars(Name), scales = "fixed") +
    labs(title = "Normalized Fm of WS blanks, outliers removed",
         x = "Normalized Fm (no LBC)")

```


# Results of the correction for dissolved secondaries

We can now use the Fm value of the water blanks to apply a large blank correction to water samples. I've used the groundwater blank for LBC since it is run more often and produces tighter data. The LBC is applied using the formulae for LBC and LBC error from SNICSer. I've looked at applying these using a per-wheel blank Fm average and a per-system average. Because each wheel with water samples didn't include two water blanks, data from the per-wheel method is sparse. The per-wheel and instrument average LBC produce very similar results, so I show only the per-system results here.


```{r}
wstd%>% 
      filter(!is.na(FmLBC),
             !is.na(fm_consensus)) %>%
  group_by(Name, rec_num, system) %>%
        summarize("No Blank" = mean(norm_ratio), 
                  "HY LBC" = mean(fm_corr),
                  "HY LBC err" = mean(sig_fm_corr),
                  "WS LBC" = mean(FmLBC),
                  "WS LBC err" = mean(FmLBCerr),
                  fm_consensus = mean(fm_consensus),
                  N = n()) %>%
        arrange(rec_num) %>%
  kable()
```

The following plot shows data for WS secondaries with no blank correction ("No LBC"), the current HY LBC ("HY LBC"), and the WS LBC as described above ("WS LBC"). Data are broken out by machine and secondary type.

```{r}
wstd %>%
  filter(!is.na(fm_consensus),
         outlier == FALSE,
         !(rec_num == 83028 & norm_ratio > 0.02)) %>%
  rename("No LBC" = norm_ratio, "HY LBC" = fm_corr, "WS LBC" = FmLBC) %>%
  pivot_longer(c("No LBC", "HY LBC", "WS LBC"), names_to = "Method", values_to = "Fm") %>%
  ggplot(aes(fct_reorder(Method, Fm, .desc = TRUE), Fm)) +
  geom_hline(aes(yintercept = fm_consensus) ) +
  geom_boxplot() +
  geom_jitter(shape = 1, color = "#0072B2") +
  facet_grid(rows = vars(fct_reorder(Name, fm_consensus)), cols = vars(system), scales = "free") +
  labs(title = "Distribution of HY and WS corrected Fm for secondary standards and blanks",
       x = "Blank cor method")
  

```

Deviation from expected value by date run, TIRI and C-2

```{r}
wstd %>%
  filter(!is.na(fm_consensus),
         outlier == FALSE,
         norm_ratio > 0.02) %>% 
  mutate(fm_norm = normFm(norm_ratio, fm_consensus)) %>%
  filter(fm_norm < 10) %>% 
  ggplot(aes(ws_strip_date, fm_norm, color = system)) +
  geom_hline(yintercept = 0) +
  geom_point()
```

WS blanks over time. 

```{r}
wstd %>%
  filter(norm_ratio < 0.015) %>% 
  ggplot(aes(ws_strip_date, norm_ratio, color = system)) +
  geom_point()
```

# Modeling the correction for varying sample Fm

Given the large blank correction formula and knowing the fraction modern of the water blank samples, we can model the effect of the correction of samples by their Fm. The simple model below assumes a WS large blank (orange line) of 4 permil and a HY blank (green line) of 1 permil. The plot shows the difference in Fm (measured - Corrected) for these two corrections. The difference between our current HY LBC and the proposed WS LBC is the difference between the orange and green lines. The difference is roughly 3 permil for dead samples, 1.6 permil for half modern, and 1/4-1/2 permil for typical Fm's of GO SHIP data.

The second plot shows the difference of secondary and blank samples from their consensus value vs Fm. The LBC models for the WS blank and HY blank are shown.


```{r}
# curve function
fun <- function(Fm, FmB) {
  bcfm <- doLBC(Fm, FmB, 1.0398)
  Fm - bcfm
}

# model plot
base <- ggplot(data.frame(Fm = c(0, 1.5)), aes(Fm))
base + 
  geom_hline(yintercept = 0) +
  stat_function(fun = fun, args = list(FmB = 0.004), aes(color = "WS LBC"), size = 1) +
  stat_function(fun = fun, args = list(FmB = 0.001), aes(color = "HY LBC"), size = 1) +
  scale_color_manual("LBC curve", values = c("orange", "green")) +
  labs(title = "Effect of LBC on Fm for HY vs WS blank",
       color = "Blank value",
       y = "Fm difference (meas - LBC)")
  
wstd %>%
  filter(!is.na(fm_consensus),
         outlier == FALSE,
         system == "USAMS",
         !(rec_num == 83028 & norm_ratio > 0.02)) %>%
  mutate(Fm = norm_ratio,
         Fm_diff = norm_ratio - fm_consensus) %>%
  ggplot(aes(Fm, Fm_diff, shape = Name)) +
  geom_hline(yintercept = 0) +
  stat_function(fun = fun, args = list(FmB = 0.004), aes(color = "WS LBC"), size = 1) +
  stat_function(fun = fun, args = list(FmB = 0.001), aes(color = "HY LBC"), size = 1) +
  geom_point() +
  xlim(0,1.5) +
  scale_color_manual("LBC curve", values = c("orange", "green")) +
  labs(title = "Data difference from consensus with HY vs WS blank curves",
       color = "Blank value",
       y = "Fm difference (meas - LBC)")
  
  
```

# Questions

Does running WS blanks and a WS correction make our data better?

If WS blanks accurately reflect the large-sample process blank and this blank does not covary with the HY large blank, the answer is yes.

Does WS blank covary with HY blank?

```{r}
data <- wstd %>% 
  filter(rec_num == 159579,
         norm_ratio < 0.01,
         lg_blk_fm < 0.006,
         total_umols_co2 * 12.015 > 1000) %>% 
  mutate(sig_norm_ratio = pmax(int_err, ext_err)) %>% 
  select(norm_ratio, sig_norm_ratio, lg_blk_fm, sig_lg_blk_fm, system)

ggplot(data, aes(norm_ratio, lg_blk_fm, color = system)) +
  geom_smooth(method = "lm") +
  geom_errorbar(aes(ymin = lg_blk_fm - sig_lg_blk_fm, ymax = lg_blk_fm + sig_lg_blk_fm)) +
  geom_errorbarh(aes(xmin = norm_ratio - sig_norm_ratio, xmax = norm_ratio + sig_norm_ratio)) +
  geom_point() +
  labs(title = "Groundwater blank Fm vs HY blank Fm",
       subtitle = "Measured on the same wheel",
       x = "Fm of groundwater blank",
       y = "Fm of hydrolysis blank")
```

This plot uses the uncorrected Fm (norm_ratio) and the Fm of the large blank used to correct the sample as paired values. Errorbars are measurement error for groundwater blanks and the large blank error for the HY blank, which is the SD of the HY blanks on a wheel.

Linear fit by system

```{r}
models <- data %>% 
  split(.$system) %>% 
  map(~ lm(lg_blk_fm ~ norm_ratio, data = .))

models %>% 
  map(summary)
```

The fit parameters confirm what the plot shows: there is a relationship between the ratio of the HY blanks and the groundwater blanks on the same wheel, but it's not particularly strong, and it explains only about 14% of the variability.

Another way to think about this is whether the HY blank is correlated with variability in the water secondaries. Working with CFAMS data only.

```{r}
data <- wstd %>% 
  filter(system == "CFAMS",
         outlier == FALSE) %>% # remove outliers
  group_by(sample_name) %>% 
  filter(n() >= 10) %>%  #only sample_names with more than 10 
  ungroup() %>% 
  mutate(sig_norm_ratio = pmax(int_err, ext_err)) %>% 
  select(sample_name, norm_ratio, sig_norm_ratio, lg_blk_fm, sig_lg_blk_fm, system, fm_consensus)
ggplot(data, aes(lg_blk_fm, norm_ratio)) +
  geom_smooth(method = "lm") +
  geom_errorbarh(aes(xmin = lg_blk_fm - sig_lg_blk_fm, xmax = lg_blk_fm + sig_lg_blk_fm)) +
  geom_errorbar(aes(ymin = norm_ratio - sig_norm_ratio, ymax = norm_ratio + sig_norm_ratio)) +
  geom_point() +
  facet_wrap(~ sample_name, scales = "free") +
  labs(title = "Groundwater blank Fm vs HY blank Fm",
       subtitle = "Measured on the same wheel",
       y = "Fm of WS sample",
       x = "Fm of hydrolysis blank")
```
```{r}
data %>% 
  nest(data = -sample_name) %>% 
  mutate(fit = map(data, ~ lm(norm_ratio ~ lg_blk_fm, data = .x)),
         tidied = map(fit, glance)) %>% 
  unnest(tidied) %>% 
  select(sample_name, adj.r.squared, statistic, p.value)

```

Or just a correlation test

```{r}
data %>% 
  nest(data = -sample_name) %>% 
  mutate(test = map(data, ~ cor.test(.x$norm_ratio, .x$lg_blk_fm)),
         tidied = map(test, tidy)) %>% 
  unnest(tidied) %>% 
  select(sample_name, estimate, p.value) 
```

There's no strong correlation between the HY blank and whether norm_ratios of secondaries are higher or lower. This may not be surprising given the variability of the blank is only a few permil, and variability of secondaries is the same or higher.

Final step is to check whether there's any correlation with variability in groundwater blank. This is harder because there's not always a groundwater blank with every wheel that has WS secondaries.

# Next steps

Using the water blanks to perform a large blank correction has a significant effect on reported ratios that differ from the ratio of the standard. This correction would improve the accuracy of water sample measurements, especially at low Fm. 

Developing a constant contamination/mass balance model for water samples would be useful for two reasons. First, samples of differing mass would get an appropriate correction. Second, large mass samples would get a correction that accounts for the fact that the normalizing standards do not contain the process blank for the water lines.