---
title: "SNICSer method comparison"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

This report compares data for a suite of wheels analyzed with the production
version of SNICSer (2.84) with results from the same wheels analyzed with a new
version of SNICSer (2.92) that includes the following changes to the blank correction
method:

1. Mass balance correction for all samples
2. Large blank average from wheel instead of long-term average for small samples
3. Total_mass instead of target_mass
4. Subtract mass of blank from total mass to get mass of unknown in MBC functions
5. New parameters for mass balance correction
6. Mass error from dc13 table
   

This method was used to analyse test wheels:

1. Load wheels in prod version of snicser
2. inherit 1st analyst's flags
3. save analysis to file
4. load wheel from file in test version
5. load saved analysis
6. blank correct
7. first authorize

The same normalization method, flagging, and large wheel blank values are used for both runs. I confirmed that norm_ratio, int_err, ext_err, lg_blk_fm, and sig_lg_blk_fm are the same for both analyses. fm_corr and sig_fm_corr are the same for all large samples, but differs for small samples due to the change in method.

```{r}
library(tidyverse)
library(here)
library(readxl)
```

## Load data


```{r}
data <- read_csv(here("data/SNICSer_blank_compare.csv"))
```

## Differences between BC methods

Differences are given as production result - test result.

### Large blank differences

Differences in the large blank corrected Fm are due to changing from long term average blank to values from the wheel. There's a small difference in in a few other wheels, but nothing greater than 1E-6. 

```{r}
lbc_diff <- data %>%
  pivot_wider(c(wheel, wheel_pos, ss, tp_num), 
              names_from = table, values_from = fm_corr) %>%
  mutate(fm_diff = snics_results - snics_results_test)

lbc_diff %>%
  filter(!near(snics_results, snics_results_test, tol = 1E-06)) %>%
  arrange(-abs(fm_diff))
```

Using large blanks from the wheel rather than stored values leads to discrepancies in fm_corr of up to {r max(lbc_diff$fm_diff)} for small samples.


### MBC Fm differences

The table below shows the differences in reported Fm between methods for all samples. Samples that did not get a MBC (primaries, Ceylon, and samples without MBC parameters) are not included. 

```{r}
data %>%
  mutate(Fmodern = ifelse(is.na(fm_mb_corr) | fm_mb_corr == 0, fm_corr, fm_mb_corr)) %>%
  pivot_wider(c(wheel, wheel_pos, tp_num, mass), 
              names_from = table, values_from = Fmodern) %>%
  mutate(fm_diff = snics_results - snics_results_test) %>%
  arrange(-abs(fm_diff))
```



## Validation of calculations 

This section compares the new SNICSer calculations to what should be the same mass balance calculations done outside of SNICSer using R and Excel (provided by Mark). Mark and I have both seen small differences before. The cause of the differences is uncertain, but they are small and do not significantly affect results.

```{r}
data %>%
  filter(table == "snics_results_test") %>%
  mutate(fm_diff = fm_mb_corr - fm_mb_corr_r) %>%
  select(wheel, wheel_pos, mass, fm_mb_corr, fm_mb_corr_r, fm_diff) %>%
  arrange(desc(fm_diff))
```


#### MR Excel data

```{r}
data %>%
  filter(table == "snics_results_test") %>%
  mutate(fm_diff = fm_mb_corr - fm_mb_corr_mr) %>%
  select(wheel, wheel_pos, mass, fm_mb_corr, fm_mb_corr_mr, fm_diff) %>%
  arrange(desc(fm_diff))
```


## Comparison with consensus

Here I'm looking at only standards that we have a known value for in the DB.


```{r}
data %>%
  filter(!is.na(normFm)) %>% 
  pivot_wider(c(wheel, wheel_pos, tp_num, mass, fm_consensus), 
              names_from = table, values_from = normFm) %>%
  mutate(norm_fm_diff = snics_results - snics_results_test) %>%
  arrange(-abs(norm_fm_diff))
```

Fm disagreement with consensus blows up at very low Fm and with some samples < 250 umol. 

```{r}
data %>%
  filter(fm_mb_corr > 0.1, total_umols_co2 < 500) %>%
ggplot(aes(total_umols_co2, normFm, color = table)) +
  geom_point()
data %>%
  #filter(Fmodern > 0.1) %>%
ggplot(aes(fm_consensus, normFm, color = table)) +
  geom_point()
data %>%
  filter(fm_mb_corr > 0.1) %>%
ggplot(aes(fm_consensus, normFm, color = table)) +
  geom_point()
```

Compare average agreement with consensus for Fm > 0.1 for new and old methods

```{r}
data %>%
  filter(fm_mb_corr > 0.1,
         !is.na(fm_consensus),
         ss == 1) %>%
  group_by(table) %>%
    summarize("Mean normalized Fm" = mean(normFm),
              "SD of norm Fm" = sd(normFm),
              "Sigma of Fm" = mean(sigma),
              "SD of Sigma" = sd(sigma),
              N = n())
```

Compare old and new, including large samples that now get MBC.

```{r}
data %>%
  filter(Fm_reported > 0.1,
         !is.na(fm_consensus)) %>%
  group_by(table) %>%
    summarize("Mean normalized Fm" = mean(normFm),
              "SD of norm Fm" = sd(normFm),
              "Sigma of Fm" = mean(sigma),
              "SD of Sigma" = sd(sigma),
              N = n())
```
